---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Data: FEMA Individual Assistance, pulled from API: https://www.fema.gov/openfema-data-page/individuals-and-households-program-valid-registrations-v1 Flood Insurance Claims, pulled from API: https://www.fema.gov/openfema-data-page/fima-nfip-redacted-claims-v1 

# Loading in packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
```

# Loading in FEMA Individual Assistance data for NC after Hurricane Matthew
```{r}
MATTHEW <- read_csv("C:/mejo390/HurricaneRebuildingNC/matthew_ihp.csv")
```

# Loading in FEMA Individual Assistance data for NC after Hurricane Florence
```{r}
FLORENCE <- read_csv("C:/mejo390/HurricaneRebuildingNC/florence_ihp.csv")
```

# Adding a column to both datasets called eligibility. Each row will either have a value of "eligible", meaning they received some form of assistance, or "ineligible", meaning they received no assistance. 
```{r}
FLORENCE <- FLORENCE %>% 
  mutate(eligibility = if_else(ihpEligible == 1 | haEligible == 1 | onaEligible == 1 | tsaEligible == 1 | rentalAssistanceEligible == 1 | repairAssistanceEligible == 1 | replacementAssistanceEligible == 1 | personalPropertyEligible == 1, "eligible", "not eligible"))

MATTHEW <- MATTHEW %>%
   mutate(eligibility = if_else(ihpEligible == 1 | haEligible == 1 | onaEligible == 1 | tsaEligible == 1 | rentalAssistanceEligible == 1 | repairAssistanceEligible == 1 | replacementAssistanceEligible == 1 | personalPropertyEligible == 1, "eligible", "not eligible"))
```

# How many applications per county? 
```{r}
MATTHEW %>%
  group_by(county) %>%
  summarise(total_m = n()) %>%
  arrange(desc(total_m))
```

```{r}
FLORENCE %>%
  group_by(county) %>%
  summarise(total_f = n()) %>%
  arrange(desc(total_f))
```

# I'm focusing on Robeson County so I'm creating datasets that filters just for that county. 
```{r}
ROBESON_F <- FLORENCE %>%
  filter(county == "Robeson (County)")

ROBESON_M <- MATTHEW %>%
  filter(county == "Robeson (County)")
```

# And now creating datasets that filters just for Lumberton, the main city in Robeson County. 
```{r}
LUMBERTON_M <- ROBESON_M %>%
  filter(damagedZipCode == 28358 | damagedZipCode == 28359 | damagedZipCode == 28360)
  
LUMBERTON_F <- ROBESON_F %>%
  filter(damagedZipCode == 28358 | damagedZipCode == 28359 | damagedZipCode == 28360)
```

# What percent of applicants in Robeson were eligible for assistance after Matthew? 
```{r}
(nrow(ROBESON_M %>% filter(eligibility == "eligible")) / nrow(ROBESON_M)) * 100
```
Just 46% of applicants in Robeson County after Hurricane Matthew were eligible for any type of individual assistance. 

# After Florence? 
```{r}
(nrow(ROBESON_F %>% filter(eligibility == "eligible")) / nrow(ROBESON_F)) * 100
```
Just 33% of applicants in Robeson County after Hurricane Florence were eligible for any type of individual assistance. 

# What percent of applicants in Lumberton were eligible for assistance after Matthew? 
```{r}
(nrow(LUMBERTON_M %>% filter(eligibility == "eligible")) / nrow(LUMBERTON_M)) * 100
```
Just 53% of applicants in Lumberton after Hurricane Matthew were eligible for any type of individual assistance. 

# After Florence? 
```{r}
(nrow(LUMBERTON_F %>% filter(eligibility == "eligible")) / nrow(LUMBERTON_F)) * 100
```
Just 38% of applicants in Lumberton after Hurricane Florence were eligible for any type of individual assistance. 

# How much money did the average eligible applicant in Robeson get after Matthew? 
```{r}
(sum(ROBESON_M$ihpAmount) + sum(ROBESON_M$rentalAssistanceAmount) + sum(ROBESON_M$replacementAmount) + sum(ROBESON_M$personalPropertyAmount)) / (nrow(ROBESON_M %>% filter(eligibility == "eligible")))
```

# After Florence? 
```{r}
(sum(ROBESON_F$ihpAmount) + sum(ROBESON_F$rentalAssistanceAmount) + sum(ROBESON_F$replacementAmount) + sum(ROBESON_F$personalPropertyAmount)) / (nrow(ROBESON_F %>% filter(eligibility == "eligible")))
```

# How much money did the average eligible applicant in Lumberton get after Matthew? 
```{r}
(sum(LUMBERTON_M$ihpAmount) + sum(LUMBERTON_M$rentalAssistanceAmount) + sum(LUMBERTON_M$replacementAmount) + sum(LUMBERTON_M$personalPropertyAmount)) / (nrow(LUMBERTON_M %>% filter(eligibility == "eligible")))
```

# After Florence? 
```{r}
(sum(LUMBERTON_F$ihpAmount) + sum(LUMBERTON_F$rentalAssistanceAmount) + sum(LUMBERTON_F$replacementAmount) + sum(LUMBERTON_F$personalPropertyAmount)) / (nrow(LUMBERTON_F %>% filter(eligibility == "eligible")))
```

# How much money did the average eligible applicant in all of Eastern NC get after Matthew? 
```{r}
(sum(MATTHEW$ihpAmount) + sum(MATTHEW$rentalAssistanceAmount) + sum(MATTHEW$replacementAmount) + sum(MATTHEW$personalPropertyAmount)) / (nrow(MATTHEW %>% filter(eligibility == "eligible")))
```

# After Florence? 
```{r}
(sum(FLORENCE$ihpAmount) + sum(FLORENCE$rentalAssistanceAmount) + sum(FLORENCE$replacementAmount) + sum(FLORENCE$personalPropertyAmount)) / (nrow(FLORENCE %>% filter(eligibility == "eligible")))
```

# How much money did the average eligible renter in Robeson get after Matthew?
```{r}
ROBESON_MR <- ROBESON_M %>%
  filter(ownRent == "Renter", eligibility == "eligible") 
  
(sum(ROBESON_MR$ihpAmount) + sum(ROBESON_MR$rentalAssistanceAmount) + sum(ROBESON_MR$replacementAmount) + sum(ROBESON_MR$personalPropertyAmount)) / nrow(ROBESON_MR)
```

# After Florence? 
```{r}
ROBESON_FR <- ROBESON_F %>%
  filter(ownRent == "Renter", eligibility == "eligible") 
  
(sum(ROBESON_FR$ihpAmount) + sum(ROBESON_FR$rentalAssistanceAmount) + sum(ROBESON_FR$replacementAmount) + sum(ROBESON_FR$personalPropertyAmount)) / nrow(ROBESON_FR)
```

# How much money did the average eligible homeowner in Robeson get after Matthew?
```{r}
ROBESON_MO <- ROBESON_M %>%
  filter(ownRent == "Owner", eligibility == "eligible") 
  
(sum(ROBESON_MO$ihpAmount) + sum(ROBESON_MO$rentalAssistanceAmount) + sum(ROBESON_MO$replacementAmount) + sum(ROBESON_MO$personalPropertyAmount)) / nrow(ROBESON_MO)
```

# After Florence? 
```{r}
ROBESON_FO <- ROBESON_F %>%
  filter(ownRent == "Owner", eligibility == "eligible") 
  
(sum(ROBESON_FO$ihpAmount) + sum(ROBESON_FO$rentalAssistanceAmount) + sum(ROBESON_FO$replacementAmount) + sum(ROBESON_FO$personalPropertyAmount)) / nrow(ROBESON_FO)
```

# Loading in flood insurance claim data
```{r}
MCLAIMS <- read_csv("C:/mejo390/HurricaneRebuildingNC/2016.csv")
FCLAIMS <- read_csv("C:/mejo390/HurricaneRebuildingNC/2018.csv")
```

# How many claims were made in Robeson after Matthew? 
```{r}
nrow(MCLAIMS %>%
  filter(countyCode == 37155))
```

# After Florence? 
```{r}
nrow(FCLAIMS %>%
  filter(countyCode == 37155))
```

# How many claims were made in Lumberton after Matthew? 
```{r}
nrow(MCLAIMS %>%
  filter(reportedZipcode == 28358 | reportedZipcode == 28359 | reportedZipcode == 28359))
```

# After Florence?
```{r}
nrow(FCLAIMS %>%
  filter(reportedZipcode == 28358 | reportedZipcode == 28359 | reportedZipcode == 28359))
```

# How much money did the average flood insurance claim in Robeson get after Matthew? 
```{r}
RMCLAIMS <- MCLAIMS %>%
  filter(countyCode == 37155)

(sum(RMCLAIMS$amountPaidOnBuildingClaim, na.rm=TRUE) + sum(RMCLAIMS$amountPaidOnContentsClaim, na.rm=TRUE) + sum(RMCLAIMS$amountPaidOnIncreasedCostOfComplianceClaim, na.rm=TRUE)) / nrow(RMCLAIMS)
```

# After Florence?
```{r}
RFCLAIMS <- FCLAIMS %>%
  filter(countyCode == 37155)

(sum(RFCLAIMS$amountPaidOnBuildingClaim, na.rm=TRUE) + sum(RFCLAIMS$amountPaidOnContentsClaim, na.rm=TRUE) + sum(RFCLAIMS$amountPaidOnIncreasedCostOfComplianceClaim, na.rm=TRUE)) / nrow(RFCLAIMS)
```

# How much money did the average flood insurance claim in Lumberton get after Matthew? 
```{r}
LMCLAIMS <- MCLAIMS %>%
   filter(reportedZipcode == 28358 | reportedZipcode == 28359 | reportedZipcode == 28359)

(sum(LMCLAIMS$amountPaidOnBuildingClaim, na.rm=TRUE) + sum(LMCLAIMS$amountPaidOnContentsClaim, na.rm=TRUE) + sum(LMCLAIMS$amountPaidOnIncreasedCostOfComplianceClaim, na.rm=TRUE)) / nrow(LMCLAIMS)
```

# After Florence?
```{r}
LFCLAIMS <- FCLAIMS %>%
   filter(reportedZipcode == 28358 | reportedZipcode == 28359 | reportedZipcode == 28359)

(sum(LFCLAIMS$amountPaidOnBuildingClaim, na.rm=TRUE) + sum(LFCLAIMS$amountPaidOnContentsClaim, na.rm=TRUE) + sum(LFCLAIMS$amountPaidOnIncreasedCostOfComplianceClaim, na.rm=TRUE)) / nrow(LFCLAIMS)
```

# Visualizing how much money an applicant can get if they have flood insurance vs. ReBuild NC vs. FEMA's individual assistance program. Using a data frame created in Excel using above calculations
```{r}
viz <- read_csv("C:/mejo390/HurricaneRebuildingNC/viz.csv")
```

```{r}
funding <- ggplot(viz, aes(fill=Program, y=Amount, x=Storms)) + 
    geom_bar(position="dodge", stat="identity") +
  labs(title = "How much money is given after a storm, and by who?", subtitle = "The average amount of money a recipient from each program received \nafter Matthew and after Florence.", caption = "Data Source: FEMA OpenData, ReBuild NC. \n Note: ReBuild NC doesn't seperate funding by storm, \nthus the average is the same for both Matthew and Florence.") +
  theme_classic() + 
  scale_fill_brewer(palette="GnBu")
```

# How percent of the county has an active flood insurance policy?
```{r}
policies <- read_csv("C:/mejo390/HurricaneRebuildingNC/policies.csv")
```

```{r}
active <- policies %>% filter(policyTerminationDate >= as.Date("2020-10-27"))
active <- active %>% filter(countyCode == 37155)
(nrow(active) / 130625) * 100
```

# How percent of the county had an active flood insurance policy during Matthew?
```{r}
activeM <- policies %>% filter(policyEffectiveDate <= as.Date("2016-10-08") & policyTerminationDate >= as.Date("2016-10-10"))
activeM <- activeM %>% filter(countyCode == 37155)
```

```{r}
activeF <- policies %>% filter(policyEffectiveDate <= as.Date("2018-09-12") &policyTerminationDate >= as.Date("2018-09-15"))
activeF <- activeF %>% filter(countyCode == 37155)
```

# What percent of each income level was eligible for assistance?
```{r}
ROBESON_M %>% 
  group_by(grossIncome, eligibility) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

ROBESON_F %>%
  group_by(grossIncome, eligibility) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

LUMBERTON_M %>% 
  group_by(grossIncome, eligibility) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

LUMBERTON_F %>% 
  group_by(grossIncome, eligibility) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```


